External Patch Group Prior Guided Internal Subspace Learning for Real Image Denoising

### 2016-10-16 egpaper_for_review.tex
a

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\mathbf{Y}=\mathbf{U}\mathbf{S}\mathbf{V}'
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component. We employ the $\mathbf{U}$ as the adaptive dictioanry to represent the structural variations of the PGs in that component.
\begin{equation}\label{equ5}
\min\nolimits_{\boldsymbol{\upalpha}}\|\mathbf{\overline{y}}_{m}-\mathbf{U}\boldsymbol{\upalpha}\|_{2}^{2}+\sum_{i=1}^{3p^{2}}\lambda_{i}|\boldsymbol{\upalpha}_{i}|.
\end{equation}
The $i$th entry of the regularization parameter $\lambda_{i}$ 
\begin{equation}\label{equ6}
\lambda_{i} = \frac{\lambda}{\mathbf{S}_{i}+\varepsilon},
\end{equation}
where $\varepsilon$ is a small positive number to avoid dividing by zero. Since the dictionary $\mathbf{U}$ is orthonormal, it is not difficult to find out that (\ref{equ4}) has a closed-form solution (detailed derivation can be found in the supplementary material):
\begin{equation}\label{equ7}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{4pt}
\hat{\boldsymbol{\upalpha}}= \text{sgn}(\mathbf{U^{T}\overline{y}}_{m})\odot \text{max}(|\mathbf{U^{T}\overline{y}}_{m}|-\mathbf{\Lambda},\mathbf{0}),
\end{equation}
where $\mathbf{\Lambda} = [\lambda_{1},\lambda_{2},...,\lambda_{3p^2}]$ is the vector of regularization parameter and $\text{sgn}(\bullet)$ is the sign function, $\odot$ means element-wise multiplication, and $|\mathbf{U^{T}\overline{y}}_{m}|$ is the absolute value of each entry of vector $|\mathbf{U^{T}\overline{y}}_{m}|$. The closed-form solution makes our weighted sparse coding process very efficient. 
\vspace{-0.05in}

\subsection{What does External Prior help the Internal Subspace Learning?}
The external data can help the internal learning in two ways. On one hand, it can guide the noisy image patches to be divided into correct subspaces through clustering. If we cluster the noisy patches in an automatical way, just like the PLE \cite{ple} did, the signal dependent noise would be hardly removed. With the help of clean external data, the noisy patches can be divided into correct subspaces. Besides, external data guided internal clustering is much more efficient than directly clustering the noisy data due to the time consuming fitting procedure. On the other hand, due to the correct division of internal noisy data, the dictionary and paranters for subspace learning could be more adaptive to the testing real noisy image. Hence, it would achieve better denoising performance than the methods only using the external information.

\subsection{The Internal Spaces Are Subspaces of External PG Spaces}
In this subsection, we compare the distribution of external PGs extracted from clean natural images and real noisy images. For better illumination, we randomly selected a cluster and project the original clean PGs onto a 2-D plane. Plot a Figure to demonstrate the data distributed in external space and the assigned internal subspace. The internal subspace is only a subspace of the external space. Hence, if we use the external data to perform denoising, the performance would be limited due to the dependence of noise on signal. If we combine the external and internal data for subspace learning, the learned subspace would be too generative and would therefore not specifically suitable for the testing data.

\section{The Enhanced Algorithm}

@@

This will also bring a second problem, that how we could automatically terminate the iteration. This can be solved by two methods. One way is to compare the images between two iterations and calculate their difference, the iteration can be terminated if the difference is smaller than a threshold. The other way is to estimate the noise level of the current image and terminate the iterations when the noise level is lower than a preset threshold. We employ the second way and set the threshold as 0.0001 in our experiments. In fact, most of our testing images will be denoised well in one iteration.

\subsection{Efficient Model Selection by Gating Network}
In the Gaussian component selection procedure, if we employ the full posterior estimation, the speed is not fast. Our algorithm can be speeded up by introducing the Gating network model.

@@

The PSNR and SSIM resluts are listed in table \ref{tab1}.
\begin{table*}\label{tab1}
\caption{Average PSNR(dB) results of different methods on 60 cropped real noisy images captured in \cite{crosschannel2016}.}
\label{tab1}
\begin{center}
\renewcommand\arraystretch{1}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 & \textbf{Noisy} &\textbf{CBM3D}&\textbf{WNNM}&\textbf{MLP}&\textbf{CSF}&\textbf{TRD}& \textbf{NI}& \textbf{NC}&\textbf{Offline} &\textbf{Online} &\textbf{Guided} 
\\
\hline
PSNR & 34.51  &  34.58 &  34.52  & 36.19   & 37.40 & 37.75 &  36.53  &  37.57  &  38.19   & 38.07 &  \textbf{ 38.51}
\\
\hline
SSIM & 0.8718  & 0.8748  & 0.8743   & 0.9470 & 0.9598 &  0.9617 & 0.9241  &  0.9514  &  0.9663   & 0.9625 &   \textbf{0.9674}
\\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}\label{tab2}
\caption{Average PSNR(dB) results of different methods on 15 cropped real noisy images used in \cite{crosschannel2016}.}
\label{tab1}
\begin{center}
\renewcommand\arraystretch{1}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
Camera Settings & \textbf{Noisy} &\textbf{CBM3D}&\textbf{WNNM}&\textbf{MLP}&\textbf{CSF}&\textbf{TRD}& \textbf{NI}& \textbf{NC}& \textbf{CC} &\textbf{Guided} 
\\
\hline
\multirow{3}{*}{\small{Canon 5D Mark III}} 
& 37.00 & 37.08 & 37.09 & 33.92 & 35.68 & 36.20 & 37.68 & 38.76 & 38.37 & \textbf{40.32}
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 33.88 & 33.94 & 33.93 & 33.24 & 34.03 & 34.35 & 34.87 & 35.69 & 35.37 & \textbf{37.12}
\\ 
\cline{2-11}    
& 33.83 & 33.88 & 33.90 & 32.37 & 32.63 & 33.10 & 34.77 & 35.54 & 34.91 & \textbf{37.09}  
\\
\hline
\multirow{3}{*}{Nikon D600} 
& 33.28 & 33.33 & 33.34 & 31.93 & 31.78 & 32.28 & 34.12 & \textbf{35.57} & 34.98 & 35.23
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 33.77 & 33.85 & 33.79 & 34.15 & 35.16 & 35.34 & 35.36 & \textbf{36.70} & 35.95 & 36.57
\\ 
\cline{2-11}    
& 34.93 & 35.02 & 34.95 & 37.89 & 39.98 & 40.51 & 38.68 & 39.28 & \textbf{41.15} & 38.73
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 35.47 & 35.54 & 35.57 & 33.77 & 34.84 & 35.09 & 37.34 & 38.01 & 37.99 & \textbf{38.52}
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 1600}   
& 35.71 & 35.79 & 35.77 & 35.89 & 38.42 & 38.65 & 38.57 & 39.05 & \textbf{40.36} & \textbf{40.36}
\\ 
\cline{2-11}    
& 34.81 & 34.92 & 34.95 & 34.25 & 35.79 & 35.85 & 37.87 & 38.20 & 38.30 & \textbf{38.81}
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 33.26 & 33.34 & 33.31 & 37.42 & 38.36 & 38.56 & 36.95 & 38.07 & \textbf{39.01} & 38.32
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 32.89 & 32.95 & 32.96 & 34.88 & 35.53 & 35.76 & 35.09 & 35.72 & \textbf{36.75} & \textbf{36.73}
\\ 
\cline{2-11}    
& 32.91 & 32.98 & 32.96 & 38.54 & 40.05 & \textbf{40.59} & 36.91 & 36.76 & 39.06 & 38.29
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 29.63 & 29.66 & 29.71 & 33.59 & 34.08 & 34.25 & 31.28 & 33.49 & \textbf{34.61} & 33.37
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 6400}   
& 29.97 & 30.01 & 29.98 & 31.55 & 32.13 & 32.38 & 31.38 & 32.79 & \textbf{33.21} & \textbf{33.23}
\\ 
\cline{2-11}    
& 29.87 & 29.90 & 29.95 & 31.42 & 31.52 & 31.76 & 31.40 & 32.86 & 33.22 & \textbf{33.29}
\\
\hline
Average PSNR & 33.41 & 33.48 & 33.48 & 34.32 & 35.33 & 35.65 & 35.49 & 36.43 & 36.88 & \textbf{ 37.07}
\\
\hline
Average SSIM & 0.8483 & 0.8511 & 0.8512 & 0.9113 & 0.9250 & 0.9280 & 0.9126 & 0.9364 & 0.9481 & \textbf{ 0.9501}
\\
\hline
\end{tabular}
\end{center}
\end{table*}

>>>>

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\mathbf{Y}=\mathbf{U}\mathbf{S}\mathbf{V}^{T}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

\subsection{What does External Prior help the Internal Subspace Learning?}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.

\subsection{Discussion on Internal Subspace Structure}
The Internal PGs are in fact lying in the subspaces of external PG Spaces. To defend this argument, we compare the distribution of external PGs extracted from clean natural images and real noisy images. For better illumination, we randomly selected a cluster and project the original clean PGs $\mathbf{X}$ onto a 2-D plane. This could be done via $\mathbf{Xp}=\mathbf{U(:.1:2)}^{T}\mathbf{X}$, where $\mathbf{U}$ is the singular vector matrix of that cluster. The noisy PGs $\mathbf{Y}$ assigned in this cluster is also projected into 2-D via $\mathbf{Yp}=\mathbf{U(:.1:2)}^{T}\mathbf{Y}$. The Figure \ref{} reflects the distribution on the 2-D plane of the projected clean PGs from external natural images and the projected noisy PGs from internal image. We can see that the internal noisy PGs are indeed lying in a subspace of the external PGs. Hence, if we directly use the external prior learned from clean PGs, the learned subspaces would be too generative to be suitable for the testing data.

Through SVD, the PGs in each internal subspace can be divided into singular vectors and singular values. The singular vectors are the basis of the corresponding subspace while the singular values reflect the importance of these basis. The basis can be used as dictionary to code the noisy PGs. And the sigular values are adaptive parameters for internal noisy PGs. We can compare the singular values of one internal subspace and the corresponding space of external PGs. The result is shown in Figure \ref{}. From which we can see that the noisy subspace often have higher values than external space consisted of clean PGs. This gap is clearly made of the noise and can be used for image denoising in a natural way.

\section{The Denoising Algorithm}
We employ the $\mathbf{U}$ as the adaptive dictioanry to represent the structural variations of the PGs in that component.
\begin{equation}\label{equ5}
\min\nolimits_{\boldsymbol{\upalpha}}\|\mathbf{\overline{y}}_{m}-\mathbf{U}\boldsymbol{\upalpha}\|_{2}^{2}+\sum_{i=1}^{3p^{2}}\lambda_{i}|\boldsymbol{\upalpha}_{i}|.
\end{equation}
The $i$th entry of the regularization parameter $\lambda_{i}$ 
\begin{equation}\label{equ6}
\lambda_{i} = \frac{\lambda}{\mathbf{S}_{i}+\varepsilon},
\end{equation}
where $\varepsilon$ is a small positive number to avoid dividing by zero. Since the dictionary $\mathbf{U}$ is orthonormal, it is not difficult to find out that (\ref{equ4}) has a closed-form solution (detailed derivation can be found in the supplementary material):
\begin{equation}\label{equ7}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{4pt}
\hat{\boldsymbol{\upalpha}}= \text{sgn}(\mathbf{U^{T}\overline{y}}_{m})\odot \text{max}(|\mathbf{U^{T}\overline{y}}_{m}|-\mathbf{\Lambda},\mathbf{0}),
\end{equation}
where $\mathbf{\Lambda} = [\lambda_{1},\lambda_{2},...,\lambda_{3p^2}]$ is the vector of regularization parameter and $\text{sgn}(\bullet)$ is the sign function, $\odot$ means element-wise multiplication, and $|\mathbf{U^{T}\overline{y}}_{m}|$ is the absolute value of each entry of vector $|\mathbf{U^{T}\overline{y}}_{m}|$. The closed-form solution makes our weighted sparse coding process very efficient.

@@

This will also bring a second problem, that how we could automatically terminate the iteration. This can be solved by two methods. One way is to compare the images between two iterations and calculate their difference, the iteration can be terminated if the difference is smaller than a threshold. The other way is to estimate the noise level of the current image and terminate the iterations when the noise level is lower than a preset threshold. We employ the second way and set the threshold as 0.002 in our experiments. In fact, most of our testing images will be denoised well in one iteration.

%\subsection{Efficient Model Selection by Gating Network}
%In the Gaussian component selection procedure, if we employ the full posterior estimation, the speed is not fast. Our algorithm can be speeded up by introducing the Gating network model.

@@

The PSNR and SSIM resluts are listed in table \ref{tab1}.
\begin{table*}\label{tab1}
\caption{Average PSNR(dB) results of different methods on 60 cropped real noisy images captured in \cite{crosschannel2016}.}
\label{tab1}
\begin{center}
\renewcommand\arraystretch{1}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 & \textbf{Noisy} &\textbf{CBM3D}&\textbf{WNNM}&\textbf{MLP}&\textbf{CSF}&\textbf{TRD}& \textbf{NI}& \textbf{NC}&\textbf{Offline} &\textbf{Online} &\textbf{Guided} 
\\
\hline
PSNR & 34.51  &  34.58 &  34.52  & 36.19   & 37.40 & 37.75 &  36.53  &  37.57  &  38.19   & 38.07 &  \textbf{ 38.55}
\\
\hline
SSIM & 0.8718  & 0.8748  & 0.8743   & 0.9470 & 0.9598 &  0.9617 & 0.9241  &  0.9514  &  0.9663   & 0.9625 &   \textbf{0.9675}
\\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}\label{tab2}
\caption{Average PSNR(dB) results of different methods on 15 cropped real noisy images used in \cite{crosschannel2016}.}
\label{tab1}
\begin{center}
\renewcommand\arraystretch{1}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
Camera Settings & \textbf{Noisy} &\textbf{CBM3D}&\textbf{WNNM}&\textbf{MLP}&\textbf{CSF}&\textbf{TRD}& \textbf{NI}& \textbf{NC}& \textbf{CC} &\textbf{Guided} 
\\
\hline
\multirow{3}{*}{\small{Canon 5D Mark III}} 
& 37.00 & 37.08 & 37.09 & 33.92 & 35.68 & 36.20 & 37.68 & 38.76 & 38.37 & \textbf{40.41}
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 33.88 & 33.94 & 33.93 & 33.24 & 34.03 & 34.35 & 34.87 & 35.69 & 35.37 & \textbf{37.14}
\\ 
\cline{2-11}    
& 33.83 & 33.88 & 33.90 & 32.37 & 32.63 & 33.10 & 34.77 & 35.54 & 34.91 & \textbf{37.11}  
\\
\hline
\multirow{3}{*}{Nikon D600} 
& 33.28 & 33.33 & 33.34 & 31.93 & 31.78 & 32.28 & 34.12 & \textbf{35.57} & 34.98 & 35.25
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 33.77 & 33.85 & 33.79 & 34.15 & 35.16 & 35.34 & 35.36 & \textbf{36.70} & 35.95 & 36.60
\\ 
\cline{2-11}    
& 34.93 & 35.02 & 34.95 & 37.89 & 39.98 & 40.51 & 38.68 & 39.28 & \textbf{41.15} & 38.78
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 35.47 & 35.54 & 35.57 & 33.77 & 34.84 & 35.09 & 37.34 & 38.01 & 37.99 & \textbf{38.59}
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 1600}   
& 35.71 & 35.79 & 35.77 & 35.89 & 38.42 & 38.65 & 38.57 & 39.05 & 40.36 & \textbf{40.44}
\\ 
\cline{2-11}    
& 34.81 & 34.92 & 34.95 & 34.25 & 35.79 & 35.85 & 37.87 & 38.20 & 38.30 & \textbf{38.85}
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 33.26 & 33.34 & 33.31 & 37.42 & 38.36 & 38.56 & 36.95 & 38.07 & \textbf{39.01} & 38.34
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 3200}   
& 32.89 & 32.95 & 32.96 & 34.88 & 35.53 & 35.76 & 35.09 & 35.72 & \textbf{36.75} & \textbf{36.74}
\\ 
\cline{2-11}    
& 32.91 & 32.98 & 32.96 & 38.54 & 40.05 & \textbf{40.59} & 36.91 & 36.76 & 39.06 & 38.29
\\
\hline
\multirow{3}{*}{Nikon D800} 
& 29.63 & 29.66 & 29.71 & 33.59 & 34.08 & 34.25 & 31.28 & 33.49 & \textbf{34.61} & 33.37
\\ 
\cline{2-11} 
\multirow{3}{*}{ISO = 6400}   
& 29.97 & 30.01 & 29.98 & 31.55 & 32.13 & 32.38 & 31.38 & 32.79 & \textbf{33.21} & \textbf{33.21}
\\ 
\cline{2-11}    
& 29.87 & 29.90 & 29.95 & 31.42 & 31.52 & 31.76 & 31.40 & 32.86 & 33.22 & \textbf{33.27}
\\
\hline
Average PSNR & 33.41 & 33.48 & 33.48 & 34.32 & 35.33 & 35.65 & 35.49 & 36.43 & 36.88 & \textbf{ 37.09}
\\
\hline
Average SSIM & 0.8483 & 0.8511 & 0.8512 & 0.9113 & 0.9250 & 0.9280 & 0.9126 & 0.9364 & 0.9481 & \textbf{ 0.9502}
\\
\hline
\end{tabular}
\end{center}
\end{table*}
